---
title: "Optimize Alpha"
output: 
    html_document:
        toc: true
        theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This package can be used to justify the alpha level based on the informational 
value of studies. The basic idea is that of a Bayesian Robot who only knows the 
alpha level, power, and whether a result was significant or not and tries to adjust
alpha and power in a way that optimizes the change in belief. The equations this is based on can be found at the end of the document. 


## idealAlpha

This function calculates the alpha level with the highest overall informational value.

```{r, echo = F}
#set working directory accordingly
source("main.R")
```

```{r, message = F, warnings = F}
#calculate best alpha level for independent sample t-test with 20 participants
#per group, expected d of 0.5 and prior of 0.5.
idealAlpha(20, 0.5, 0.5)
```
The most informative alpha level, the power, and the expected correct change in 
belief are returned. The expected correct change in belief is defined as subtracting
the probability-weighted updating after a false positive or false negative from the
weighted updating after a true positive or true negative. 

In addition, a plot shows informational value as a function of alpha level, a blue 
line indicates the optimal alpha and the red line the 5% alpha
level. 

## alphaConfirmation

This function allows to supply an intended probability after a significant result. 
It will select the most informative alpha level under which the intended posterior 
probability after a significant result will still be achieved. 

```{r, message = F, warnings = F}
#calculate the alpha level needed for a four group anova with cohens f = 0.4,
#ten participants per group, a prior of 0.5 and an intended posterior probability of 0.95
alphaConfirmation(20, 0.4, 0.5, 0.95, test = "anova")
```


## alphaDisconfirmation

This function works similar to alphaConfirmation but targets false negatives. 
In other words, it allows to supply the intended probability after a 
non-significant finding and will select the most informative alpha level under
which the intended posterior probability after a non-significant result will still be achieved. 

```{r, message = F, warnings = F}
#Caluclate the alpha level needed for a one sided within subjects t-test, with 40 participants, 
#d = 0.5, prior of 0.5 and intended posterior probability of 0.1 after a non-significant result. 
alphaDisconfirmation(40, 0.3, 0.5, 0.1, type = "one.sample", alternative = "greater")
```

## probPower

Combines alphaConfirmation and alphaDisconfirmation to a probability based power
analysis. Fix both the intended posterior probability after a significant result 
as well as a non-significant result, the function will return alpha, power and 
sample size needed to achieve this. 

```{r, message = F, warnings = F}
#calculate the alpha, power and sample size needed to achieve 20% posterior probability after a non significant result and 95% posterior probability after a significant result with a prior of 0.5 and an effect size of 0.2 for between samples t-test. 
probPower(0.2, 0.5, 0.2, 0.95)
```

## postConfirm and postDisconfirm

PostConfirm calculates the posterior probability after a significant result for a given alpha, power, and prior and postDisconfirm calculates the posterior probability after a non-significant result for a given power, alpha, and prior. 

```{r, message = F, warnings = F}
#calculate posterior probability after significant and non-significant results for alpha = 0.05, power = 0.8, and prior = 0.5
postConfirm(0.8, 0.05, 0.5)
postDisconfirm(0.8, 0.05, 0.5)
```

## Equations

We can calculate the posterior probability after a significant result using Bayes rule as:

\begin{equation}
p(H_1/+) = \frac{(1-\beta) \times p(H_1)}{(1-\beta) \times p(H_1)+ \alpha \times (1-p(H_1))}
\label{eq:confirmation}
\end{equation}

and for a non-significant result as: 

\begin{equation}
p(H_1/-) = \frac{\beta \times p(H_1)}{\beta \times p(H_1)+(1-\alpha) \times (1-p(H_1))}
\label{eq:disconfirmation}
\end{equation}

Combining the two, the expected correct change in belief for the idealAlpha function can be calculated as follows:

\begin{equation}
\label{eq:L}
\begin{aligned}
\delta L = p(H_1) \times (1-\beta) \times (p(H_1/+)-p(H_1))+ \\ (1-p(H_1)) \times (1-\alpha) \times (p(H_1)-p(H_1/-)) - \\ (1-p(H_1)) \times \alpha \times (p(H_1/+)- p(H_1))- \\ p(H_1) \times \beta \times (p(H_1)-p(H_1/-))
\end{aligned}
\end{equation}
